---
title: "Stats 506 F20, Group Project"
subtitle: "ARIMA Model Tutorial"
author: "Group 3: Zhilin He, Jialun Li, Chuwen Li"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    theme: cerulean
---

```{r setup, include=FALSE}
# 79: -------------------------------------------------------------------------
#! We generally don't need to see the code in the results document.
knitr::opts_chunk$set(echo = FALSE)
#! Make sure all chunks have a label. This one is labelled 'setup'.
```

## Introduction

ARIMA models are good. (This is a stub.)


## Data Description

We use the [nifty fifty stocks data]
(https://www.kaggle.com/rohanrao/nifty50-stock-market-data) from NSE India as 
an example in this tutorial.

```{r dataset, include=FALSE}
nifty = read.csv('./NIFTY50_all.csv')
```

The data contains daily open, close, highest and lowest prices, volume and 
other relevant information for the "Nifty Fifty" stocks since January 2000. 
There are 50 stocks, 15 variables and 230104 observations in total. Detailed 
variable descriptions are shown in Table 1 below.

| Variables Name   |Variable Description     |
| ------------- |:-------------:|
| Date   | Date               |
| Symbol | Symbol of the stock (name, identifier)  |
| Series | Type of security (“EQ” for all rows)   |
| Prev Close | Previous day's close price |
| Open | Open price of day |
| High | Highest price in day |
| Low | Lowest price in day |
| Last | Last traded price in day |
| Close | Close price of day |
| VWAP | Volume Weighted Average Price |
| Volume | Volume |
| Turnover | Turnover |
| Trades | Number of Trades (Half missing) |
| Deliverable Volume | Amount of deliverable volume (7\% missing) |
| \%Deliverable | Percent of deliverable volume (7\% missing) |
<center>
**Table 1**. Variables in the dataset
</center>

Each variable other than <code>Date</code>, <code>Symbol</code> and 
<code>Series</code> is a time series starting from January 3rd, 2000 to
September 30th, 2020. Missing data is a problem in this dataset, so we try to
deal with missing data first. We can use ARIMA models to model these time
series data. We are going to show how to build ARIMA models in Python, R and
Stata respectively. 

As we are more interested in the stock prices, we use the variable
<code>VWAP</code> for the most part. It can summarize the average price of the
stock on a trading day. We want to catch the trend of stock prices across the
years and possibly forecast future stock prices.

## Core Example {.tabset .tabset-pills}

### Python

```{python echo=FALSE}
```

### R

### Stata

## To-do list


<!-- ## Summary -->

## References

1. A modern Time Series tutorial:
[Link](https://www.kaggle.com/rohanrao/a-modern-time-series-tutorial)

2. ARIMA model in Wikipedia:
[Link](https://en.wikipedia.org/wiki/Autoregressive_integrated_moving_average)