---
title: "Stats 506 F20, Group Project"
subtitle: "ARIMA models for Time-Series Data"
author: "Group 3: Zhilin He, Jialun Li, Chuwen Li"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---

-----------

```{r , echo = FALSE, warning = FALSE, message = FALSE}
# this block loads R packages that may be needed for the analysis.
library(dplyr)
library(ggplot2)
```

```{r , echo = FALSE, warning = FALSE, message = FALSE}
# load data
url = 'https://raw.githubusercontent.com/lixx4228/Stats506_group3/main'
nifty = readr::read_delim(sprintf('%s/NIFTY50_all.csv', url), delim = ',')

```

# Introduction

This project is a tutorial of using multivaraite time series analysis for the 
stock market index, NIFTY 50 from NSE (National Stock Exchange) India. The 
data is obtained from 
[Nifty 50](https://www.kaggle.com/rohanrao/nifty50-stock-market-data) 
contains price history and trading volumes of fifty stocks in India from 
2000-01-03 to 2020-09-30. 

We illustrates how to using Python, R, and Stata to apply Auto Regressive 
Integrated Moving Average (ARIMA) to time series data. ARIMA is able to fit 
a given non-seasonal non-stationary time series based on its lag values.
A general ARIMA model consists of three parts: the "AR" part means the variable
of interest is regressed on its lag terms, the "I" part means the differenced
values are used, and the "MA" part means the regression error is modelled as a
linear combination of error terms in the past. The purpose of using differenced
terms is to make the time series stationary for autoregression.

An ARIMA model is characterized by 3 terms: p (the order of AR term), q 
(the order of the MA term), and d (number of differencing to make time series 
stationary). Given a time series \(\{X_t\}\), an \(ARIMA(p, d, q)\) model 
can be expressed as: 
$$(1-\sum_{i=1}^p\phi_iL^i)(1-L)^dX_t=
(1+\sum_{i=1}^q\theta_iL^i)\epsilon_t + \delta$$
where \(\epsilon_t\) is the error term, \(L\) is the lag operator, i.e.
\(LX_t = X_{t-1}, \forall t>1\), \(p\) is the number of lagged terms of \(X\),
\(d\) is the number of times of differencing needed for stationarity,
\(q\) is the number of lagged forecast errors in prediction, \(\delta\)
is the interception term for the regression, and \(\theta, \phi\)'s are 
the estimated regression coefficients.

ARIMA models are fitted in order to understand the data better and forecast
future data. They are based on linear regression models. The best model can
be chosen using AIC or BIC.

In Section 2 We cleaned the dataset and conducted basic data exploratory 
analysis to see whether there is non-stationarity (trend or seasonality) in 
the time series prior to applying the ARIMA models. Section 3 has three 
subsections, which illustrates using Python, R, and Stata for analysis 
respectively. Section 4 gives the summary results and conclusion of this 
tutorial.

# Data Summary

`NIFTY 50` data consist of 50 stocks, 230104 observations on 15 variables. 
The data contains daily open, close, highest and lowest prices, volume and 
other relevant information for the "Nifty Fifty" stocks since January 2000. 
Detailed variable descriptions are shown in Table 1 below.


```{r , echo = FALSE, warning = FALSE, message = FALSE}
# variable description
var_names = names(nifty)
var_names[16] = "Return"
tab1 = data.frame(name = var_names, 
                  description = 
                    c("Date of trade", "Name of the company",
                      "We have only one series: Equity(EQ)",
                     "Previous day's close price", "Open price of day",
                     "Highest price in day", "Lowest price in day",
                     "Last traded price in day", "Close price of day",
                     "Volume Weighted Average Price",
                     "A measure of sellers versus buyers of a particular stock", 
                     "The number of shares available for sale",
                     "The number of shares traded", 
                  "Shares which are actually transferred among demat accounts",
                  "Percent of deliverble volume",
                  "Return of trade"))

```

<details>
<summary> Click to see variable descriptions. </summary>
```{r var_tab, echo = FALSE}
tab.cap1 = '**Table 1**. *Varaibel descriptions of NIFTY 50 data*'
col_names = c("Variable Name", "Variable Description")
tab1 %>%
  knitr::kable(format = 'html',  col.names = col_names, caption = tab.cap1) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) %>% 
  kableExtra::column_spec(1, italic = T) 

```
</details>

As we are more interested in the stock prices, we use the variable
<code>VWAP</code> for the most part. It can summarize the average price of the
stock on a trading day. We want to catch the trend of stock prices across the
years and possibly forecast future stock prices.

```{r clean_data, warning = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)

old_sym = c('MUNDRAPORT', 'UTIBANK', 'BAJAUTOFIN', 'BHARTI', 'HEROHONDA',
       'HINDALC0', 'HINDLEVER', 'INFOSYSTCH', 'JSWSTL', 'KOTAKMAH', 'TELCO',
       'TISCO', 'UNIPHOS', 'SESAGOA', 'SSLT', 'ZEETELE')
new_sym = c('ADANIPORTS', 'AXISBANK', 'BAJFINANCE', 'BHARTIARTL', 'HEROMOTOCO',
       'HINDALCO', 'HINDUNILVR', 'INFY', 'JSWSTEEL', 'KOTAKBANK', 'TATAMOTORS',
       'TATASTEEL', 'UPL', 'VEDL', 'VEDL', 'ZEEL')

# summary statistics of variable of interest
nifty = nifty %>%
  mutate(Symbol = lapply(Symbol, 
                         function(x) replace(x, x %in% old_sym, new_sym)),
         Trades = ifelse(is.na(Trades), 0, Trades),
         Return = Close - `Prev Close`) %>% 
  select(Date, Symbol, Return, VWAP, Volume, Trades)
summary(nifty[, -2])
```

```{r trend, warning = FALSE, fig.cap = fig.cap1, fig.width = 6.8}
# plot trend of all stocks
fig.cap1 = "**Figure 1.** *Daily trend of all stocks, 2000-2020.*"

nifty_ts = reshape2::melt(nifty[, -c(2)], id.vars = "Date")
ggplot(nifty_ts, aes(x = Date, y = value)) + 
    geom_line(color="lightblue") + 
    theme_bw() +
    facet_wrap(~ variable, scales = "free_y", ncol = 1)

```

From the trend of all stocks, we can see the time series exhibit 
non-stationarity. There was a substantial strike to the India stock market 
after the outbreak of Coronavirus.

# Time Series Analysis {.tabset .tabset-fade .tabset-pills}

## Python

### Data cleaning and visualization


## R

### Autocorrelation Plots

We will focus on the variables: `Symbol`, `VWAP`, `Volume`, `Trades` and a 
newly created variable `Return`, which is the difference between `Close` and 
`Prev Close`. First we choose one stock to analyze its autocorrelation of 
trend on theabove four variables. The trend plot above shows large value of 
`Volume`, sowe can take log to decrease its trend.

```{r acf, warning = FALSE, message = FALSE, fig.cap = fig.cap2}
library(forecast)

candidate = "ADANIPORTS"

nifty_cand = nifty %>% filter(Symbol == candidate) %>% 
  mutate(Volume = log(Volume))

vars_list = names(nifty_cand)[-c(1, 2)]
fig.cap2 = "**Figure 2.** *ACF plots of stock: ADANIPORTS.*"

par(mfrow=c(2,2))
for ( var in vars_list ) {
  df_ts = ts(nifty_cand[[var]], frequency = 1, start = c(2000, 01, 03))
  # acf
  acf_tou <- acf(df_ts, lag.max = 30, plot = FALSE)
  plot(acf_tou, xlab = "Lag (in Year)", main = var)
}

```

Notice that all the variables show high autocorrelation except for `Return`, 
which is because `Return` is calculated from the first difference of closing 
price working as a linear filter applied to eliminate a trend. Since we are 
going to apply ARIMA model to the data, which can works for non-stationary 
time series, we can leave the model to detrend time seires. For the 
illustration purpose, we will take first difference of other three variables 
and compare the autocorrelation plot to the previous one.

```{r , echo=FALSE, warning=FALSE, message=FALSE}
acf_diff_plot = function(df = nifty, sym = "ADANIPORTS") {
  stopifnot( length(sym) == 1)
  stopifnot( sym %in% df$Symbol)
  
  df = df %>% filter(Symbol == sym) %>% 
    mutate(Volume = log(Volume))

  vars_list = names(df)[-c(1, 2)]
  par(mfrow=c(2,2))
  for ( var in vars_list ) {
    if (var != "Return"){
    df_ts = ts(df[[var]], frequency = 1, start = c(2000, 01, 03))
    df_ts = diff(df_ts, 1)
    } else {
    df_ts = ts(df[[var]], frequency = 1, start = c(2000, 01, 03))
    }
    # acf
    acf_tou <- acf(df_ts, lag.max = 30, plot = FALSE)
    plot(acf_tou, xlab = "Lag (in Year)", main = var)
  }
} 
```

```{r acf2, warning=FALSE, message=FALSE, fig.cap = fig.cap3}
fig.cap3 = "**Figure 3.** *ACF plots of first difference of time seires.*"
acf_diff_plot(df = nifty, sym = candidate)
```

### Fitting an ARIMA Model

 We can choose the ARIMA Model by AIC. Let’s tabulate some AIC values for 
 a range of different choices of p and q, assuming d takes 0 for `Return` 
 while 1 for other 3 variables. Below shows the AIC table of fitting ARIMA 
 on `Return` time series of stock: "ADANIPORTS". We will subset the last 120 
 time series as test data.
 
```{r aic, echo=FALSE, warning = FALSE,message = FALSE}
aic_table <- function(ts, P, Q, d){ 
  table <- matrix(NA, (P + 1), (Q + 1)) 
  for(p in 0:P) { 
    for(q in 0:Q) { 
      table[p + 1, q + 1] <- arima(ts, order=c(p, d, q))$aic
    } 
  }
  dimnames(table) <- list(paste("AR", 0:P, sep = ""), 
                          paste("MA", 0:Q, sep = ""))
  table
}

# Construct AIC table
nifty_cand_ts = ts(nifty_cand[, -2]$Return, 
                   frequency = 1, start = c(2000, 01, 03))
nifty_aic_table <- aic_table(head(nifty_cand_ts, -30), 4, 4, 0) 

tab.cap2 = '**Table 2**. *AIC for different ARIMA parameters*'
nifty_aic_table %>%
  knitr::kable(format = 'html', caption = tab.cap2) %>%
  kableExtra::kable_styling('striped', full_width = TRUE) 


```

The AIC table suggests that ARIMA(4, 0, 4) is the best model for the return of 
"ADANIPORTS". This model may give a sign that increasing p and q will tend to 
get smaller AIC for a better fit. However, models with higher p and q are 
more complex, so it may lead to problems like overfitting, numerical stability 
and etc. We usually prefer a simply model, which also better for interpretation.

Even though it is nice to view the change of AIC value as the change of 
p and q, for a big dataset like this, it is very inefficient to iterate over 
range of p and q. `auto.arima()`in the `forest` package is much faster in 
generating the results. We can verify that `auto.arima()` also suggests 
ARIMA(4, 0, 0) is the model with the best fit in the range of (1, 4) of p and q.

```{r , warning=FALSE, message=FALSE}
ts_arima <- auto.arima(head(nifty_cand_ts, -30), max.p = 4, 
                      max.q = 4, max.d = 3)
print(ts_arima)
```

### Forecasting using an ARIMA Model

Lastly, we will forecast the next 120 time series and compare the result with 
our test set.
```{r acc, warning=FALSE, message=FALSE}
ts_forecasts <- forecast(ts_arima, h = 30) 
acc <- accuracy(ts_forecasts, head(tail(nifty_cand_ts, 30), 7))
print(round(acc, 4))
```
The RMSE and MAE for the test set are 5.5276 and 4.5046, respectively.

```{r forc_fig, warning=FALSE, fig.cap = fig.cap4, fig.height=4, fig.width=6.8}
fig.cap4 = "**Figure 4.** *Forecasts from ARIMA(4, 0, 4) with zero mean.*"
autoplot(ts_forecasts, main = "") + xlab("Day") + 
  ylab("Return") +
  theme_bw()
```

## Stata

# Conclusion


# References

1. A modern Time Series tutorial:
[Link](https://www.kaggle.com/rohanrao/a-modern-time-series-tutorial)

2. ARIMA model in Wikipedia:
[Link](https://en.wikipedia.org/wiki/Autoregressive_integrated_moving_average)





